Estimating Access to Rabies Post-Exposure Prophylaxis in Madagascar

### 2019-03-18 accessPEP.Rmd
with mapping of added pts

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
## Clear memory
rm(list = ls())
## Libraries
library(data.table)
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
library(stringr)
library(ISOweek)
library(raster)
library(sp)
library(MASS)
library(reshape2)
library(geosphere)
library(foreach)
library(bbmle)
library(boot)
library(geosphere)

@@

## Get color palette from other network fig!
ctar_metadata$color <- c("#FCC56F","#FFDBE5", "#7A4900", "#CBCDD2", "#0000A6", "#EFF2F1",
                 "#99d8c9", "#B79762", "#004D43", "#8FB0FF", "#997D87", "#FD9C54", "#8362B5",
                 "#578FB0","#5A0007", "#809693", "#D16100", "#1B4400", "#4FC601", "#3B5DFF", 
                 "#4A3B53", "#FF2F80","#61615A", "#BA0900", "#6B7900", "#00C2A0", "#FFAA92",
                 "#FF90C9", "#B903AA", "#FEFFE6", "#E9E9D2")
ctar_metadata$color[ctar_metadata$CTAR %in% no_data] <- "#D3D3D3"
ctar_metadata$fill <- ctar_metadata$color
ctar_metadata$fill <- add.alpha(ctar_metadata$fill, 0.4)
# display_palette(ctar_metadata$color)

@@

## Getting district data
mada_district@data %>% 
  select(mdg_dis_co, district, pop = pop2015un, ttimes_weighted, ctar_in_dist, 
         dist_catch = ctar_dist, exclude, color, fill) -> dist_data

exps_dist <- get.district.data(ctar_data = ctar_data, ctar_metadata = ctar_metadata, 
                               mada_district = mada_district, catchments, 
                               consec_days = 15, 
                               contact_cutoff = 3, reporting_thresh = 0.25, 
                               start = ymd("2014-01-01"), end = ymd("2017-12-31"))
exps_mora <- get.morabites(mada_communes = mada_communes)

## Getting other data
rep_data <- get.reporting(ctar_data, ctar_metadata, start_date = ymd("2014-01-01"), 
                          end_date = ymd("2017-12-31"), 
                          threshold = 10)
ctar_contacts <- get.contacts(ctar_data, times_sd = 3)
ctar_contacts <- ctar_contacts[["ctar_data"]]
```

@@

### Bite patient data
We used a database of bite patient forms submitted to IPM from ARMC across the country between 2014 - 2017. These were individual patient data forms that were submitted as frequently as monthly to annually by clinics. Of the 31 existing ARMC, `r length(no_data) - 2` submitted fewer than 10 forms over the four years. Two clinics, the IPM ARMC and the Fianarantsoa ARMC had separate databases which were not available at the time of analysis. Overall, we had data from *`r 31 - length(no_data)`* clinics across the country (Fig 1A). These data include details of the administrative district of the bite patient's address and the date of reporting. We also had *27* months of data from the Moramanga District that were resolved to the commune level (the administrative level below the district).

@@

@@

We calculated the expected reporting of rabid exposures ($\overline\rho$) given bite incidence as predicted by our model($\mu$) as:
$$
  \overline\rho = \frac{\mu \times p_{rabid}}{R}
$$
or the fraction of incidence due to rabid exposures ($\mu \times p_{rabid}$) divided by the total rabies exposure incidence ($r$) for a range of estimated rabies incidence and $p_{rabid}$. We look at the range of $p_{rabid}$ reported in Rajeev et al. 2018 for data from the Moramanga District (0.2 - 0.6). where the proportion of reported bites that are rabies exposures ($p_{rabid}$) are defined as:
$$
           p_{rabid}=
            \begin{cases}
            x, & \text{if}\ \frac{R \times \rho_{max}}{B} > x \\
            \frac{R \times \rho_{max}}{B}, & \text{otherwise}
            \end{cases}
$$
<br>

@@

### Sensitivity analyses for burden estimates 
We estimate burden deterministically across a range of parameter values to test the effects of our model assumptions on estimates of rabies burden. Specifically, we fix rabies incidence at the minimum and maximum of our estimated range, look at the range of values of p_rabid between 0.2 - 0.6 (as per Moramanga), and the range of values of rho_max (0.85 - 0.99) to get at maximum and minimum estimates of burden. We also examine the impact of systematic variation in rabies incidence with human population (a potential proxy for changes in the dog population/HDR) by looking at how estimates of burden change if rabies incidence were to scale positively or negatively with human population.

@@

## profile CIs
prof_CI <- lapply(mada_mods, confint)
bite_mods <- as.data.frame(list(beta = mapply(function(x) coef(x)["beta"], mada_mods),
                                intercept = mapply(function(x) coef(x)["intercept"], mada_mods),
                                beta_upper = mapply(function(x) x["beta", 2], prof_CI),
                                beta_lower = mapply(function(x) x["beta", 1], prof_CI),
                                intercept_lower = mapply(function(x) x["intercept", 1], prof_CI),
                                intercept_upper = mapply(function(x) x["intercept", 2], prof_CI),
                                likelihood = mapply(function(x) logLik(x), mada_mods),
                                model = mapply(function(x) attributes(x)$label, mada_mods)))
## Predictions: fixed
ttimes_plot <- seq(0, 25, by = 0.5)
distance_plot <- seq(0, 200, by = 2.5)
pop_plot <- seq(1000, 1e6, length.out = 30)

library(iterators)
preds_fixed <- run.predicts.fixed(covar_df = bite_mods, pop_fixed = 1e5, pop_seq = pop_plot)
preds_data <- run.predicts.data(covar_df = bite_mods, run_type = "predict.data")
preds_raw <- run.predicts.data(covar_df = bite_mods, run_type = "predict.raw")

library(patchwork)
a <- ggplot(preds_fixed, aes(x = covar, y = predicted, color = scale)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = scale), color = NA, alpha = 0.5) +
  scale_color_manual(values = c("navy", "blue", "purple")) +
  scale_fill_manual(values = c("blue", "lightblue", "pink"), name = "Model scale") +
  xlab("Travel time (hrs)") +
  ylab("Reported bites per \n 100,000 persons") +
  labs(tag = "A")

@@

The best fitting model did not include population as a predictor for the district data...write up from supplement!

@@

ggplot(rep_min_max, aes(x = ttimes, y = rep_min, fill = scale)) +
  geom_line(aes(y = rep_min, color = scale), linetype = 2) +
  geom_line(aes(y = rep_max, color = scale), linetype = 1) +
  geom_ribbon(aes(ymax = rep_max, ymin = rep_min), alpha = 0.5) +
  scale_fill_manual(values = c("navy", "blue")) +
  scale_color_manual(values = c("navy", "blue"))

@@

system.time({
  results_comm_med <- get.burden.stochastic(names = comm_mat$mdg_cm_, ttimes = comm_mat$ttimes/60,
                           pop = comm_mat$pop, scenario = comm_mat$scenario,
                           param_ttimes = B_ttimes_mada,
                           param_intercept = B_0_mada,
                           p_rab_min = 0.2, p_rab_max = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000)
})

@@

## Baseline results
results_all %>% filter(scenario == 0) -> results_all_base
ggplot(results_all_base, aes(y = reorder(district, ttimes_weighted), x = deaths_mean, 
                              color = model)) +
  ggstance::geom_pointrangeh(aes(xmin = deaths_lowerCI, xmax = deaths_upperCI), fatten = 2) +
  ylab("") +
  xlab("Average annual deaths") +
  scale_color_manual(values = c("district" = alpha("blue", 0.5), 
                                "commune" = alpha("purple", 0.5)), 
                     name = "Scale") -> p
p + theme(axis.text.y = element_text(margin = margin(t = 10, r = 0, b = 10, l = 0), size = 7))

## Deaths averted
ggplot(results_all_base, aes(y = reorder(district, ttimes_weighted), x = averted_mean, 
                              color = model)) +
  ggstance::geom_pointrangeh(aes(xmin = averted_lowerCI, xmax = averted_upperCI), fatten = 2) +
  ylab("") +
  xlab("Average annual deaths") +
  scale_color_manual(values = c("district" = alpha("blue", 0.5), 
                                "commune" = alpha("purple", 0.5)), 
                     name = "Scale") -> p
p + theme(axis.text.y = element_text(margin = margin(t = 10, r = 0, b = 10, l = 0), size = 7))

@@

``` {r scenario}
## Plotting national level
ggplot(data = results_natl, aes (x = scenario, y = deaths_mean, color = model)) +
  geom_line() +
  geom_ribbon(aes(ymin = deaths_upperCI, ymax = deaths_lowerCI, 
                  color = NULL, fill = model),
              alpha = 0.5) +
  scale_color_manual(values = c("district" = "blue", 
                                "commune" = "purple"), 
                     guide = "none") +
  scale_fill_manual(values = c("district" = "blue", 
                                "commune" = "purple")) +
  guides(color = guide_legend(), 
         fill = guide_legend(override.aes = list(linetype = 1)))

@@

## Deaths averted
ggplot(data = results_natl, aes (x = scenario, y = averted_mean, color = model)) +
  geom_line() +
  geom_ribbon(aes(ymin = averted_upperCI, ymax = averted_lowerCI, 
                  color = NULL, fill = model),
              alpha = 0.5) +
  scale_color_manual(values = c("district" = "blue", "commune" = "purple"),  guide = "none") +
  scale_fill_manual(values = c("district" = "blue", "commune" = "purple")) +
  guides(color = guide_legend(), fill = guide_legend(override.aes = list(linetype = 1)))

@@

```{r vial analysis}
# ## new ctar
rep_data <- get.reporting(ctar_data, ctar_metadata, start_date = ymd("2014-01-01"), 
                          end_date = ymd("2017-12-31"), 
                          threshold = 10)
date_mat <- rep_data[["date_mat"]]
dose_mat <- rep_data[["doses_wide"]]
colnames(date_mat) <- colnames(dose_mat)[2:(ncol(dose_mat)-1)]
doses <- melt(dose_mat, id.vars = "date_de_consultation", value.name = "patients", variable.name = "id_ctar")

@@

ggplot(compare_dists, aes(x = patients, y = prop_plot, fill = type)) +
  geom_col(color = NA, alpha = 0.5, position = "identity") +
  scale_fill_manual(values = c("red", "blue")) +
  facet_wrap(~id_ctar, scales = "free") +
  xlab("# of Patients") +
  ylab("Proportion of days")

rep_bites = seq(0, 15000, by = 100)
predicts <- as.data.frame(list(rep_bites = rep_bites,
                               predicted_mu = exp(0.029*rep_bites/100 + -0.307),
                               predicted_k = exp(0.022*rep_bites/100 + -0.39)))
ggplot(predicts, aes(x = rep_bites, y = predicted_mu)) +
  geom_line()
ggplot(predicts, aes(x = rep_bites, y = predicted_k)) +
  geom_line()

@@

### Key findings

### Strengths and Limitations

### Broader context

### Conclusions

@@

## Supplementary Materials
### S1. RLE thresholds for reporting

### S2. Removing contacts from the data set

@@

ggpairs(data = select(exps_mora, pop, ttimes_weighted, distance))

@@

kable(AICctab(mada_mods, nobs = nrow(exps_dist), mnames = mapply(function(x) attributes(x)$label, mada_mods), weights = TRUE))

mora_mods <- c(mora_mods_covars, mora_mods_pop)
kable(AICctab(mora_mods, nobs = nrow(exps_mora), mnames = mapply(function(x) attributes(x)$label, mora_mods), weights = TRUE))

@@

### S4. Sensitivity of model parameter estimates to contact and RLE cut-offs

>>>>

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
## Clear memory
rm(list = ls())
## Libraries
library(data.table)
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
library(stringr)
library(ISOweek)
library(raster)
library(sp)
library(MASS)
library(reshape2)
library(geosphere)
library(foreach)
library(bbmle)
library(boot)
library(geosphere)
library(iterators)
library(patchwork)

@@

## Get color palette from other network fig!
ctar_metadata$color <- c("#FCC56F","#FFDBE5", "#7A4900", "#CBCDD2", "#0000A6", "#EFF2F1",
                 "#99d8c9", "#B79762", "#004D43", "#8FB0FF", "#997D87", "#FD9C54", "#8362B5",
                 "#578FB0","#5A0007", "#809693", "#D16100", "#1B4400", "#4FC601", "#3B5DFF", 
                 "#4A3B53", "#FF2F80","#61615A", "#BA0900", "#6B7900", "#00C2A0", "#FFAA92",
                 "#FF90C9", "#B903AA", "#FEFFE6", "#E9E9D2")
ctar_metadata$fill <- ctar_metadata$color
ctar_metadata$color[ctar_metadata$CTAR %in% no_data] <- "#D3D3D3"
# ctar_metadata$fill <- add.alpha(ctar_metadata$fill, 0.4)
# display_palette(ctar_metadata$color)

@@

## Getting district covars
district_coords <- coordinates(mada_district)
distance_mat <- distm(district_coords, ctar_coords)
mada_district$dist_min <- apply(distance_mat, 1, min)/1000
mada_district@data %>%
  select(district = mdg_dis_co, dist_name = district, pop = pop2015adj, 
         ttimes_weighted_dist = ttimes_weighted, 
         dist_catch = CTAR, distance = dist_min, exclude, color) -> dist_data

exps <- get.district.data(ctar_data = ctar_data, covars = dist_data, ctar_metadata, 
                               consec_days = 15, 
                               contact_cutoff = 3, reporting_thresh = 0.25, 
                               start = ymd("2014-01-01"), end = ymd("2017-12-31"))
exps %>%
  group_by(district) %>%
  summarise(bites = mean(bites), n = n()) %>%
  left_join(dist_data) -> exps_dist

exps_mora <- get.morabites(mada_communes = mada_communes)
exps_mora <- exps_mora[["mora_bites"]]
## Other data
ctar_contacts <- get.contacts(ctar_data, times_sd = 3)
ctar_contacts <- ctar_contacts[["ctar_data"]]

@@

### Bite patient data
We used a database of bite patient forms submitted to IPM from ARMC across the country between 2014 - 2017. These were individual patient data forms that were submitted as frequently as monthly to annually by clinics. Of the 31 existing ARMC, `r length(no_data) - 2` submitted fewer than 10 forms over the four years. Two clinics, the IPM ARMC and the Fianarantsoa ARMC had separate databases which were not available at the time of analysis. Overall, we had data from `r 31 - length(no_data)` clinics across the country (Fig 1A). These data include details of the administrative district of the bite patient's address and the date of reporting. We also had 27 months of data from the Moramanga District that were resolved to the commune level (the administrative level below the district).

@@

```{r raw data, cache = TRUE}
## Network fig (make with ggplot!)
## Revise: points = CTAR points with size proportional to number of bites reported, lines to district proportional to the number of bites from that district, outline = by district, communes = polygons shaded by catchments

## get exposure matrix = first set of points
## district exposures
ctar_data %>% 
  drop_na(id_ctar, district) %>%
  group_by(district) %>% # group by district
  summarize(n = n()) -> exps_district # get count of exposures
## ctar exposures
ctar_data %>% 
  drop_na(id_ctar) %>%
  group_by(id_ctar) %>% # group by ctar
  summarize(n = n()) -> exps_ctar
## ctar by district exposure matrix
ctar_data %>% 
  drop_na(id_ctar, district) %>%
  group_by(district, id_ctar) %>% # group by district and ctar
  summarize(n = n()) -> exp_mat# get count of exposures

## To plot district bites
dist_points <- cbind(select(mada_district@data,
                            mdg_dis_co), coordinates(mada_district))
names(dist_points)[2:3] <- c("dist_long", "dist_lat")
dist_points %>%
  left_join(select(ctar_metadata, ctar_dist, LATITUDE, LONGITUDE), 
            by = c("mdg_dis_co" = "ctar_dist")) %>%
  mutate(dist_long = coalesce(LONGITUDE, dist_long), 
         dist_lat = coalesce(LATITUDE, dist_lat)) -> dist_points

exps_district %>%
  right_join(select(dist_points, mdg_dis_co, dist_long, dist_lat),
             by = c("district" = "mdg_dis_co")) %>%
  left_join(dist_catchments, by = c("district" = "mdg_dis_co") ) -> dist_points
dist_points$size <- log(dist_points$n + 0.1)*0.25

## To plot ctar points (only those not excluded)
exps_ctar %>%
  left_join(select(ctar_metadata, id_ctar, LATITUDE, LONGITUDE, 
                   exclude, color, fill)) %>%
  filter(exclude == 0) -> ctar_points

## To plot lines between ctar + districts
exp_mat %>% 
  left_join(select(dist_points, district, dist_long, dist_lat)) %>%
  left_join(select(ctar_metadata, id_ctar, LATITUDE, LONGITUDE,
                   color, fill, ctar_dist)) -> dist_lines
dist_lines$size <- log(dist_lines$n + 0.1)*0.25

gg_district <- fortify(mada_district, region = "mdg_dis_co")
gg_district %>% 
  left_join(mada_district@data, by = c("id" = "mdg_dis_co")) -> gg_district

cols <- c("#FCC56F","#FFDBE5", "#7A4900", "#CBCDD2", "#0000A6", "#EFF2F1",
                         "#99d8c9", "#B79762", "#004D43", "#8FB0FF", "#997D87", "#FD9C54", "#8362B5",
                         "#578FB0","#5A0007", "#809693", "#D16100", "#1B4400", "#4FC601", "#3B5DFF", 
                         "#4A3B53", "#FF2F80","#61615A", "#BA0900", "#6B7900", "#00C2A0", "#FFAA92",
                         "#FF90C9", "#B903AA", "#FEFFE6", "#E9E9D2", "#D3D3D3")
names(cols) <- as.character(cols)

sizes <- log(c(100, 200, 400, 800, 1600) + 0.1)
p <- ggplot() +
  geom_polygon(data = gg_district, aes(x = long, y = lat, group = group, fill = color), color = "grey50", alpha = 0.5) +
  geom_point(data = ctar_points, aes(x = LONGITUDE, y = LATITUDE, size = log(n + 0.1), color = color), shape = 1) +
  geom_point(data = dist_points, aes(x = dist_long, y = dist_lat, color = fill, size = log(n + 0.1)*0.75)) +
  geom_segment(data = dist_lines, aes(x = dist_long, y = dist_lat, xend = LONGITUDE, yend = LATITUDE, 
                                      color = color, size = size*1.5), alpha = 0.75, show.legend = FALSE) +
  scale_fill_manual(values = cols, guide = "none") +
  scale_color_manual(values = cols, guide = "none") +
  scale_size_identity("Reported bites", labels = as.character(c(100, 200, 400, 800, 1600)),
                      breaks = sizes, guide = "legend") +
  guides(size = guide_legend(override.aes = list(color = "grey50"))) +
  labs(tag = "A") +
  xlab("") +
  ylab("") +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank())

### Part 2
col_vals <- dist_catchments$color
names(col_vals) <- dist_catchments$mdg_dis_co

c <- ggplot(exps, aes(x = reorder(dist_name, -ttimes_weighted_dist), y = bites/pop*1e5, color = district)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  scale_color_manual(values = col_vals) + 
  ylab("Annual bites per 100k") +
  xlab("Districts ordered by \n travel times (high to low)") +
  coord_flip() +
  theme(axis.text.y = element_blank()) +
  labs(tag = "B")

p + plot_spacer() + c + plot_layout(nrow = 1, widths = c(1, 0.1, 1))

@@

We calculated the expected reporting of rabid exposures ($\overline\rho$) given bite incidence as predicted by our model($\mu$) as:
$$
  \overline\rho = \frac{\mu p_{rabid}}{R}
$$
or the fraction of incidence due to rabid exposures ($\mu \times p_{rabid}$) divided by the total rabies exposure incidence ($r$) for a range of estimated rabies incidence and $p_{rabid}$. We look at the range of $p_{rabid}$ reported in Rajeev et al. 2018 for data from the Moramanga District (0.2 - 0.6). where the proportion of reported bites that are rabies exposures ($p_{rabid}$) are defined as:
$$
           p_{rabid}=
            \begin{cases}
            x, & \text{if}\ \frac{R\rho_{max}}{B} > x \\
            \frac{R\rho_{max}}{B}, & \text{otherwise}
            \end{cases}
$$
<br>

@@

### Sensitivity analyses for burden estimates 
We estimate burden deterministically across a range of parameter values to test the effects of our model assumptions on estimates of rabies burden. Specifically, we fix rabies incidence at the minimum and maximum of our estimated range, look at the range of values of p_rabid between 0.2 - 0.6 (as per Moramanga), and the range of values of rho_max (0.85 - 0.99) to get at maximum and minimum estimates of burden. We also examine the impact of systematic variation in rabies incidence with human population (a potential proxy for changes in the dog population/HDR, see SN) by looking at how estimates of burden change if rabies incidence were to scale positively or negatively with human population.

@@

## profile CIs
prof_CI <- lapply(mada_mods, confint)
bite_mods <- as.data.frame(list(beta = mapply(function(x) coef(x)["beta"], mada_mods),
                                intercept = mapply(function(x) coef(x)["intercept"], mada_mods),
                                beta_upper = mapply(function(x) x["beta", 2], prof_CI),
                                beta_lower = mapply(function(x) x["beta", 1], prof_CI),
                                intercept_lower = mapply(function(x) x["intercept", 1], prof_CI),
                                intercept_upper = mapply(function(x) x["intercept", 2], prof_CI),
                                likelihood = mapply(function(x) logLik(x), mada_mods),
                                model = mapply(function(x) attributes(x)$label, mada_mods)))
kable(bite_mods)
## Predictions: fixed
ttimes_plot <- seq(0, 25, by = 0.5)
distance_plot <- seq(0, 200, by = 2.5)
pop_plot <- seq(1000, 1e6, length.out = 30)

preds_fixed <- run.predicts.fixed(covar_df = bite_mods, pop_fixed = 1e5, pop_seq = pop_plot)
preds_data <- run.predicts.data(covar_df = bite_mods, run_type = "predict.data")
preds_raw <- run.predicts.data(covar_df = bite_mods, run_type = "predict.raw")

a <- ggplot(preds_fixed, aes(x = covar, y = predicted, color = scale)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = scale), color = NA, alpha = 0.5) +
  scale_color_manual(values = c("navy", "blue", "purple"), guide = "none") +
  scale_fill_manual(values = c("blue", "lightblue", "pink"), name = "Model scale") +
  xlab("Travel time (hrs)") +
  ylab("Reported bites per \n 100,000 persons") +
  labs(tag = "A")

@@

@@

c <- ggplot(rep_min_max, aes(x = ttimes, y = rep_min, fill = scale)) +
  geom_line(aes(y = rep_min, color = scale), linetype = 2) +
  geom_line(aes(y = rep_max, color = scale), linetype = 1) +
  geom_ribbon(aes(ymax = rep_max, ymin = rep_min), alpha = 0.5) +
  scale_fill_manual(values = c("navy", "purple")) +
  scale_color_manual(values = c("navy", "purple")) +
  xlab("Reporting") +
  ylab("Travel times (in hours)") +
  labs(tag = "A")

@@

results_comm_med <- get.burden.stochastic(names = comm_mat$mdg_cm_, ttimes = comm_mat$ttimes/60,
                           pop = comm_mat$pop, scenario = comm_mat$scenario,
                           param_ttimes = B_ttimes_mada,
                           param_intercept = B_0_mada,
                           p_rab_min = 0.2, p_rab_max = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000)

@@

## Baseline results
results_all %>% filter(scenario == 0) -> results_all_base
ggplot(filter(results_all_base, !is.na(district)), 
       aes(y = deaths_mean, x = reorder(district, ttimes_weighted), 
                              color = model)) +
  geom_pointrange(aes(ymin = deaths_lowerCI, ymax = deaths_upperCI), fatten = 2) +
  xlab("Districts ordered by travel times \n (low to high)") +
  ylab("Average annual deaths") +
  scale_color_manual(values = c("district" = alpha("blue", 0.5), 
                                "commune" = alpha("purple", 0.5)), 
                     name = "Scale") +
  theme(axis.text.x = element_blank()) +
  labs(tag = "B") -> d

c / d

## Deaths averted
ggplot(filter(results_all_base, !is.na(district)), 
       aes(y = reorder(district, ttimes_weighted), x = averted_mean, 
                              color = model)) +
  ggstance::geom_pointrangeh(aes(xmin = averted_lowerCI, xmax = averted_upperCI), fatten = 2) +
  ylab("") +
  xlab("Average annual deaths averted") +
  scale_color_manual(values = c("district" = alpha("blue", 0.5), 
                                "commune" = alpha("purple", 0.5)), 
                     name = "Scale") +
  theme(axis.text.y = element_text(margin = margin(t = 10, r = 0, b = 10, l = 0), size = 7))

@@

``` {r scenario, cache = TRUE}
##
ctar_pts <- read.csv("output/incremental_ARMC_nofilter.csv")
ctar_pts$order <- 1:nrow(ctar_pts)
ggplot() + geom_polygon(data = gg_district, aes(x = long, y = lat, group = group), fill = "grey",
                         color = "grey50", alpha = 0.5) +
  geom_point(data = ctar_metadata, aes(x = LONGITUDE, y = LATITUDE), color = "black", shape = 1) +
  geom_point(data = ctar_pts, aes(x = Y_COORD, y = X_COORD, color = order, alpha = order))

## Plotting national level
ggplot(data = results_natl, aes (x = scenario, y = deaths_mean, color = model)) +
  geom_line() +
  geom_ribbon(aes(ymin = deaths_upperCI, ymax = deaths_lowerCI, 
                  color = NULL, fill = model),
              alpha = 0.5) +
  scale_color_manual(values = c("district" = "blue", 
                                "commune" = "purple"), 
                     guide = "none") +
  scale_fill_manual(values = c("district" = "blue", 
                                "commune" = "purple")) +
  guides(color = guide_legend(), 
         fill = guide_legend(override.aes = list(linetype = 1))) +
  xlab("# of Additional ARMC") +
  ylab("Mean annual deaths \n due to rabies") +
  labs(tag = "A") -> e

@@

## Deaths averted
ggplot(data = results_natl, aes (x = scenario, y = averted_mean*0.16, color = model)) +
  geom_line() +
  geom_ribbon(aes(ymin = averted_upperCI, ymax = averted_lowerCI, 
                  color = NULL, fill = model),
              alpha = 0.5) +
  scale_color_manual(values = c("district" = "blue", "commune" = "purple"),  guide = "none") +
  scale_fill_manual(values = c("district" = "blue", "commune" = "purple"), guide = "none") +
  xlab("# of Additional ARMC") +
  ylab("Mean annual \n deaths averted") +
  labs(tag = "B") -> f

e / f

@@

```{r vial analysis, cache = TRUE}
# ## new ctar
rep_data <- get.reporting(ctar_data, ctar_metadata, start_date = ymd("2014-01-01"), 
                          end_date = ymd("2017-12-31"), 
                          threshold = 10)
date_mat <- rep_data[["check_mat"]]
dose_mat <- rep_data[["doses_wide"]]
colnames(date_mat) <- colnames(dose_mat)[2:(ncol(dose_mat)-1)]
doses <- melt(dose_mat, id.vars = "date_de_consultation", value.name = "patients", variable.name = "id_ctar")

@@

ggplot(compare_dists, aes(x = patients, y = prop_plot, fill = type)) +
  geom_col(color = NA, alpha = 0.5, position = "identity") +
  scale_fill_manual(values = c("red", "blue")) +
  facet_wrap(~id_ctar, scales = "free") +
  xlab("# of Patients") +
  ylab("Proportion of days") -> a

rep_bites = seq(0, 15000, by = 100)
predicts <- as.data.frame(list(rep_bites = rep_bites,
                               predicted_mu = exp(0.029*rep_bites/100 + -0.307),
                               predicted_k = exp(0.022*rep_bites/100 + -0.39)))
ggplot(predicts, aes(x = rep_bites)) +
  geom_line(aes(y = predicted_mu), color = "red") +
  geom_line(aes(y = predicted_k), color = "blue") +
  xlab("Annual reported bites to a clinic") +
  ylab("Predicted parameter") -> b

a / b

@@

Main findings
- [ ] Travel times are a good predictor of bites, suggesting that this could shape burden significantly
- [ ] We develop a framework to look at the potential impact of this access on burden, to get more realistic sub-national and national estimates
- [ ] Scenario analysis shows that expanding access could significantly reduce burden
- [ ] However due to broad remoteness, expanding access alone will not eliminate human rabies deaths

Limitations
- [ ] Not thinking about clinic functioning
- [ ] Reporting issues and what it means (i.e. we're assuming that clinics are not reporting, but they may not be provisioning at all, etc.) 
- [ ] Not accounting for other factors which drive access/reporting (i.e. socioeconomics and awareness although these are likely correlated and that has implications for the scenario analysis)
- [ ] Just another set of assumptions and a set that needs validation--what data could do this?
- [ ] Contacts issue and what it means for the results
- [ ] Approach only works in endemic setting, and where reporting = receiving (i.e. in a free setting), although for GAVI countries this is likely the case

Discussion: broader context
- [ ] Compare to previous estimates of burden (Moramanga, GAVI paper, etc.) for Mada
- [ ] Compare range of estimates of human rabies exposure incidence and deaths to estimates from other endemic settings
- [ ] P_rabid estimates compared to other endemic settings (look at Moramanga data in Supplementary materials?) Differences in reporting of rabid vs. non-rabid bites?
- [ ] Availability of PEP in other countries is also similarly or even more limited
- [ ] Other studies that have looked at access and mortality/vacc

Discussion: conclusions
- [ ] Access likely shapes burden in a big way in Mada
- [ ] Expanding access will therefore likely decrease burden in a big way, but there are limitations due to infrastructure, etc.
- [ ] Spatial differences in reporting also has implications for surveillance (and could parallel dog vaccination heterogeneity) and this needs to be explored *hint*hint*
- [ ] Our results depend on strong assumptions, so key data on dog pops, rabies, reporting, etc. are needed to validate this, surveillance @ clinic level!
- [ ] Regardless expanding access alone will likely not prevent all deaths, you need dog vacc!

@@

## Supplementary Materials

### S1. Estimates of reporting
```{r}
## Getting other data
rep_data <- get.reporting(ctar_data, ctar_metadata, start_date = ymd("2014-01-01"), 
                          end_date = ymd("2017-12-31"), 
                          threshold = 15)
clinic_reporting <- rep_data[["clinic_reporting"]]
patient_ts <- rep_data[["patient_ts"]]
date_mat <- rep_data[["date_mat"]]
dose_mat <- rep_data[["dose_mat"]]
doses_wide <- rep_data[["doses_wide"]]

patient_ts %>%
  ungroup() %>%
  mutate(id_ctar = as.factor(id_ctar)) %>%
  right_join(date_mat) %>%
  mutate(n = replace_na(n, 0), 
         n = ifelse(exclude == 1, n, NA)) -> doses_long
doses_long$ctar <- ctar_metadata$CTAR[match(doses_long$id_ctar, ctar_metadata$id_ctar)]
doses_long %>%
  mutate(year = year(date_de_consultation)) %>%
  left_join(select(clinic_reporting, ctar, prop, year = date_de_consultation)) -> doses_long

p <- ggplot(data = doses_long, aes(x = date_de_consultation, y = reorder(ctar, prop))) + 
  geom_tile(aes(fill = n)) +
  scale_fill_gradientn(colours = c("white", "purple", "black"), values = c(0, 0.1, 1),
                       na.value = alpha("lightgrey", 1), name = "Number of \n patients") +
  xlim(ymd("2014-01-01"), ymd("2017-12-31")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Year") +
  ylab("ARMC") +
  labs(tag = "A")

col_vals <- ctar_metadata$fill
names(col_vals) <- ctar_metadata$CTAR

n <- clinic_reporting %>%
  mutate(prop15 = prop, prop5 = get.days(doses_wide = doses_wide, threshold = 5)[[1]]$prop,
         prop30 = get.days(doses_wide = doses_wide, threshold = 30)[[1]]$prop) %>%
  ggplot(., aes(x = reorder(ctar, prop), color = ctar)) +
  geom_hline(yintercept = 0.25, color = "darkgrey", alpha = 0.5) +
  geom_boxplot(aes(ymin = prop30, lower = prop15, middle = prop15, upper = prop15, ymax = prop5),
               stat = "identity") +
  scale_color_manual(values = col_vals) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "none") +
  xlab("") +
  ylab("Proportion of days included") +
  coord_flip() +
  labs(tag = "B")

p + n

### S2. Accounting for low-risk contacts reporting to clinics

```{r excluding contacts}
ctar_contacts <- get.contacts(ctar_data, times_sd = 3)
patient_ts <- ctar_contacts[["patient_ts"]]
patient_ts <- patient_ts[patient_ts$ctar != "Tsiroanomandidy", ]
A <- ggplot(data = drop_na(patient_ts, exclude), aes(x = ctar, y = n, color = as.factor(exclude))) +
  scale_color_manual(name = "Exclude", values = c("navy", "red"), labels = c("No", "Yes")) +
  ylab("Number of patients per day") +
  xlab("ARMC") +
  geom_jitter(alpha = 0.5, width = 0.25) +
  coord_flip() +
  labs(tag = "A")

## Testing with Moramanga data 
ctar <- get.morabites(mada_communes = mada_communes)
ctar <- ctar[["ctar"]]
sdev <- seq(2, 10, by = 0.1)
check <- data.frame(sd = NA, prop_bites = NA, prop_contacts = NA)
for (i in 1:length(sdev)){
  ctar %>%
    group_by(date.reported) %>%
    summarize(n = n()) %>%
    mutate(exclude = ifelse(n > mean(n) + sdev[i]*sd(n), 1, 0)) -> mora_ts
  ctar$exclude <-mora_ts$exclude[match(ctar$date.reported, mora_ts$date.reported)]
  checksd <- table(ctar$consult.type, ctar$exclude)[, 2]/table(ctar$consult.type)
  check <- rbind(check, c(sdev[i], checksd[1:2]))
}
check %>% drop_na() %>% gather(prop_bites, prop_contacts, -sd) -> check
B <- ggplot(check, aes(x = sd, y = prop_contacts, color = prop_bites)) +
  geom_line() +
  geom_line() +
  scale_color_manual(name = "Type of patient", values = c("blue", "red"), 
                     labels = c("Bites/scratches", "Low-risk contacts")) + 
  xlab("Number of standard \n deviations above the mean") +
  ylab("Proportion excluded") +
  labs(tag = "B")

A + B

@@

@@

print(AICctab(mada_mods, nobs = nrow(exps_dist), mnames = mapply(function(x) attributes(x)$label, mada_mods), weights = TRUE))

mora_mods <- c(mora_mods_covars, mora_mods_pop)
print(AICctab(mora_mods, nobs = nrow(exps_mora), mnames = mapply(function(x) attributes(x)$label, mora_mods), weights = TRUE))

@@

### S4. Sensitivity of parameter estimation to different reporting estimates and exclusion of contacts
